\part{The Robot Coworker} % Main chapter title

\label{part:robot_coworker} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\lhead{Part 2. \emph{The Robot Coworker}} % Change X to a consecutive number; this is for the header on each page - perhaps a

Robots can be used to perform a large number of different operations. Some of these will be simple enough that the robot can just achieve its task by performing a prefixed sequence of elementary actions. In other cases, the robot might have to achieve complex goals, which require the ability to create plans and to adapt them to the current state of the world. When cooperating with other agents, the robot has to build a shared plan, which includes the actions that every agent need to perform, in order to coordinate and ensure the corrent achievement of the goal. We can imagine the following process:
\begin{itemize}
	\item The system receives a new goal. This can be directly introduced by a human, or chosen after some kind of reasoning by the robot.
	\item One of the agents (the robot or the humans) proposes a plan to achieve the goal, and presents it to the other agents.
	\item The agents negotiate the plan. In some situations, one of the agents might not be able (or might not want) to perform a specific action, or sequence of actions. The agent can refuse the plan, proposing a correction or a completely different plan.
	\item The agents execute the plan. Each agent performs its part of the plan. In addition, agents may check the state of others to monitor the correct execution of their part of the plan or to cordinate with them.
	\item An agent might fail its part of the plan. If this happens the agents need to create a new plan to account for this failure.
 	\item The process continues until the goal is achieved or it becomes unachievable (for example, a needed resource is no longer available).
\end{itemize}

When humans cooperate this process can be very quick. For simple tasks humans are able to coordinate without explicitly forming a plan, in particular if they are used to cooperating together. Other times, when there are unexpected problems during the execution of a plan, humans are able to quickly readapt their plan, without completely restarting this process. In order to cooperate in a natural way with humans, robots need to reproduce these mechanisms.

In this part we will introduce the mechanisms that we developed for the robot coworker problem. In this problem, the robot has to achieve a goal together with a human, cooperating to solve the task. The main aspects of our works are the management of plans, shown in chapter~\ref{chapter:plan_management}, and their execution, particularly the execution of joint actions, shown in chapter ~\ref{chapter-task_execution}. Chapter ~\ref{chapter-coworker_experiments} shows our experiments in robot coworker problem. We also show, in chapter ~\ref{chapter-madp} a recent extension to this work, involving multi-agent probabilistic planning.

\input{parts/coworker/plan_management.tex} 
\input{parts/coworker/task_execution.tex} 
\input{parts/coworker/coworker_experiments.tex} 
\input{parts/coworker/mamdp.tex} 
