\part{The Robot Teacher} % Main chapter title

\label{part:robot_teacher} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\lhead{Part IV. \emph{The Robot Teacher}} % Change X to a consecutive number; this is for the header on each page - perhaps a

We can imagine robots in several ways. The more realistic persons will imagine the robot as a tool, able to execute repetitive, high-precision tasks, reducing humans' effort. The more imaginative persons will instead think of the kind of robots that we can see in science fiction literature or movies: autonomous agents, able to reason and to act almost as humans. In general, we would think of the robot as some sort of `programmable' entity. We teach the robot, through our code and algorithms, to execute its task.

In the last years, actually, there is an increasing interest in `inverting this role' by having robots act as teachers, particularly with children. Some studies, like \cite{movellan2009sociable}, have shown that a robot teacher can actually positively impact on the learning process of children.

In this part, we introduce the problem of the \textit{robot teacher}, where the robot needs to explain a user how to execute a task and then guide him in its achievement. We are particularly interested in adapting this process to the \textit{knowledge} of users in the tasks to perform, by explaining more in-depth processes where the user has less \textit{expertise}, and quickly tasks that user has already accomplished several times. While we are calling this problem \textit{robot teacher}, these mechanisms can benefit any cooperative application where the robot is acting as leader, and needs to explain a cooperative plan to its partners.  

Chapter~\ref{chapter:knowledge} will explain how we model the knowledge of a user in a task to adapt the plan explanation and management processes. Chapter~\ref{chapter:teacher_results} shows a user study performed to validate our approach.

\input{parts/teacher/human_knowledge.tex} 
\input{parts/teacher/teacher_results.tex} 
